---
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

::: {style="text-align: center; color: #337DFF;"}
<!-- --Your name, date and assignment  -->
```         
Bridget Lang
9/21/2024
Assignment 1
```

#Load relevant software libraries
```{r}
library(dplyr)
library(tidyverse)
```
#Load the data:
https://github.com/wellsdata/CompText_Jour/raw/main/data/blackindex_master.csvLinks to an external site. -->
```{r}
data <- read.csv("https://raw.githubusercontent.com/wellsdata/CompText_Jour/main/data/blackindex_master.csv")

```

#Using code, describe the number of rows and columns in the dataset
```{r}
nrow(data)
ncol(data)
#1803 rows, 30 columns
```
#Create a table that displays a count of the top 5 newspaper_city entries
```{r}
head(
  data %>%
  filter(newspaper_city != 'N/A') %>%
  count(newspaper_city) %>%
  arrange(desc(n)),
  5)

```
#Create a table that counts the different entries in the "newspaper_city" column
```{r}
  data %>%
  count(newspaper_city) 
```

#Create a table that counts all newspaper_city entries by year. UPDATE: Filter out the NA values
```{r}

years_filtered <- data %>%
  filter(newspaper_city != 'N/A') %>%
  group_by(year) %>%
  count(newspaper_city)

# data %>%
#   group_by(year) %>%
#   count(newspaper_city) %>%
#   filter(newspaper_city != 'N/A')
```

#Create a simple column ggplot chart that shows the total entries by year

```{r}

years_unfiltered <- data %>%
  group_by(year) %>%
  count(newspaper_city)

ggplot(years_unfiltered, aes(x=year, y=n)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
#This is your title sequence
  labs(title = "Total Articles by Year",
       caption = "Graphic by Bridget Lang, 9-21-2024",
       y="Number of Articles",
       x="Year")


```

Armed with your data analysis, write a 250 word memo at the end of the R markdown document that summarizes your findings. Use this opportunity to ask questions or note any major flaws in the data.

This data shows a variety of articles from different newspapers from different cities, published between the years 1890 and 2002. There are precisely 1803 articles referenced in this data, with 30 columns of information stored for each article. The majority of articles were published by newspapers in Pittsburgh, Chicago, and Atlanta, as well as Norfolk and New York. 587 out of the 714 articles whose cities were documented were published in these five cities, which is about 82%. Given the common use of the word "lynching" in the articles' titles, which is apparent just by observation, and the time frame in which the majority of articles were published, I would say that these articles discuss violence against black people, primarily lynching. Based on the above graphic, the majority of articles were published between 1880 and 1950.

One issue I noticed with this data is that a majority (1089) of the articles did not have any record of the city of publication, which limits the ability to analyze this data from a geographical or cultural standpoint. Another potential flaw in this data is duplicate and/or unused columns, which makes the data clunky to look at in a table. 

I think it would be very interesting to be able to do a textual analysis of the contents of these articles. I also think it would be interesting to look at the "document type" column, and see any variations in content or place of publication between entries marked as "front page" vs those marked as "article". 
