---
title: 'Week 2: Assignment 1'
author: "Rob Wells"
date: "2024-08-22"
output: html_document
---
# Create an R markdown document with the following:
# Your name, date and assignment 

# Create a new code chunk: Option + Apple + i
# Insert a pipe command: Shift + Apple + m

#Load relevant software libraries
```{r}
library(tidyverse)
```

#Load the mob_analysis_data.csv
```{r}
mob <- read.csv("~/Code/CompText_Jour/data/mob_analysis_data.csv")


```


#Using code, describe the number of rows and columns in the dataset
```{r}
ncol(mob)
nrow(mob)
#Dimensions: Gives number rows, then columns
dim(mob)
```

#Create a table that displays a count of the top 15 newspaper entries
```{r}
mob %>% 
  count(Newspaper) %>% 
  arrange(desc(n))

```


#Create a table that counts the different entries in the "keyword" column
```{r}
mob %>% 
  count(keyword) %>% 
  arrange(desc(n))


```

#Create a table that counts all entries by year. Call it "years"
```{r}
years <- mob %>% 
  count(Year) %>% 
  arrange(Year)


```

#Create a simple column ggplot chart that shows the total entries by year
```{r}
ggplot(data=years) +
  geom_col(mapping=aes(x=Year, y=n))


```

Armed with your data analysis, write a 250 word memo at the end of the R markdown document that summarizes your findings. Use this opportunity to ask questions or note any major flaws in the data.
