---
title: "R proficiency homework"
author: "Wells"
date: "2025-02-06"
output: html_document
---

```{r}
library(tidyverse)
```

--Load the mob analysis dataLinks to an external site.
```{r}
mob <- read.csv("https://raw.githubusercontent.com/wellsdata/CompText_Jour/main/data/mob_analysis_data.csv")
```

Links to an external site.--Using code, describe the number of rows in the dataset
```{r}
nrow(mob)
```
columns in the dataset
```{r}
ncol(mob)
```

A more sophisticated method using print and paste0 that integrates text and r output

```{r}
print(paste0("This dataset has ", nrow(mob), " rows and ", ncol(mob), " columns."))

```


--Create a table that displays a count of the top 15 newspaper entries
top_n function
```{r}

top <- mob |> 
  #filter nas
  filter(!is.na(Newspaper)) |> 
  count(Newspaper) |>
  top_n(15, n) |> 
  arrange(desc(n))

top

```

--Create a table that counts the different entries in the "keyword" column
```{r}
keyword <- mob |> 
  count(keyword) |>
  arrange(desc(n))

```

--Create a table that counts all entries by year

```{r}
year <- mob |> 
  count(Year) |> 
  arrange(Year)

year
```

--Create a simple column ggplot chart that shows the total entries by year

```{r}
mob |> 
  count(Year) |> 
   ggplot(aes(x = Year, y = n,
             fill = n)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
#This is your title sequence
  labs(title = "Count of mob appearing in newspaper articles",
       subtitle = "Lynching news research",
       caption = "Graphic by Rob Wells, 2-6-2024",
       y="Count of mob terms",
       x="Year")

```
#question from Tiasia: figure out how to space out my x-axis so that the years can be seen clearly to better understand my graph. 
```{r}
year_count <- mob_analysis |>
  count(Year)

#her version
ggplot(year_count, aes(x = factor(Year), y = n)) +
  geom_col(fill = "steelblue") +
  labs(title = "Total Entries by Year",
       x = "Year",
       y = "Total Entries") +
  theme_minimal()
```

issue with factor(Year)
ggplot(year_count, aes(x = factor(Year), y = n)) +
you use factor(Year) when you want each year to have its own distinct bar in a bar plot or you're only dealing with specific years rather than a continuous range

Don't factor(Year) and keep it numeric to control the spacing of years (like using scale_x_continuous). factor turns a variable into a categorical variable. You can't do math on categorical variables


```{r}

ggplot(year_count, aes(x = Year, y = n)) +
  geom_col(fill = "steelblue") +
  labs(title = "Total Entries by Year",
       x = "Year",
       y = "Total Entries") +
  theme_minimal() +
    theme(axis.text.x = element_text(angle=90)) +
  scale_x_continuous(breaks = seq(1800, 1960, by = 10))
```



