---
title: "Sentimentr Example"
author: "Wells"
date: "2026-10-06"
output: html_document
---

Sentimentr is a package that examines sentiment at the sentence level. 


Here is the supporting documentation
https://github.com/trinker/sentimentr

sentimentr is designed to quickly calculate text polarity sentiment in the English language at the sentence level and optionally aggregate by rows or grouping variable(s).

sentimentr is a response to my own needs with sentiment detection that were not addressed by the current R tools. My own polarity function in the qdap package is slower on larger data sets. It is a dictionary lookup approach that tries to incorporate weighting for valence shifters (negation and amplifiers/deamplifiers). Matthew Jockers created the syuzhet package that utilizes dictionary lookups for the Bing, NRC, and Afinn methods as well as a custom dictionary. He also utilizes a wrapper for the Stanford coreNLP which uses much more sophisticated analysis. Jocker's dictionary methods are fast but are more prone to error in the case of valence shifters. Jocker's addressed these critiques explaining that the method is good with regard to analyzing general sentiment in a piece of literature. He points to the accuracy of the Stanford detection as well. In my own work I need better accuracy than a simple dictionary lookup; something that considers valence shifters yet optimizes speed which the Stanford's parser does not. This leads to a trade off of speed vs. accuracy. Simply, sentimentr attempts to balance accuracy and speed.

Why sentimentr
So what does sentimentr do that other packages don't and why does it matter?

sentimentr attempts to take into account valence shifters (i.e., negators, amplifiers (intensifiers), de-amplifiers (downtoners), and adversative conjunctions) while maintaining speed. Simply put, sentimentr is an augmented dictionary lookup. The next questions address why it matters.

So what are these valence shifters?

A negator flips the sign of a polarized word (e.g., "I do not like it."). See lexicon::hash_valence_shifters[y==1] for examples. An amplifier (intensifier) increases the impact of a polarized word (e.g., "I really like it."). See lexicon::hash_valence_shifters[y==2] for examples. A de-amplifier (downtoner) reduces the impact of a polarized word (e.g., "I hardly like it."). See lexicon::hash_valence_shifters[y==3] for examples. An adversative conjunction overrules the previous clause containing a polarized word (e.g., "I like it but it's not worth it."). See lexicon::hash_valence_shifters[y==4] for examples.


```{r}
#install.packages("sentimentr")
library(tidyverse)
library(sentimentr)
library(tidyr)
```



# Example dataframe

```{r}
reviews <- tibble(
  id = 1:2,
  text = c(
    "The food was not good, but the service was extremely friendly.",
    "I barely enjoyed the movie, although the actors were talented."
  )
)
```

# 1. Split into sentences and calculate sentiment at the sentence level
```{r}
sentence_sentiment <- reviews %>%
  get_sentences() %>%
  sentiment_by(by = c("id")) # Summarizes back to row/document level if desired
```



# View sentence-by-sentence detailed breakdown
```{r}
detailed_sentences <- reviews %>%
  get_sentences() %>%
  sentiment()

head(detailed_sentences)
```

#Trump speech
```{r}

#trump <- read_tsv("trump_dallas_sept_9_2026.txt")
trump <- read.table("trump_dallas_sept_9_2026.txt", header = F, sep = "\t") |> 
  rename(text = V1) |> 
  mutate(text = str_replace(text, "Trump:", "")) |> 
  mutate(id = row_number())

```


# 1. Split into sentences and calculate sentiment at the sentence level
```{r}
trump_sentiment <- trump %>%
  get_sentences() %>%
  sentiment_by(by = c("id")) 
```

#What is the most negative sentence?

```{r}


trump$text[89]
```

