---
title: "Homework_Assignment"
author: "Taufiq"
date: "2024-09-17"
output: html_document
---

```{r}
#install the libraries 
library(janitor)
library(tidyverse)
library(rio)
```

```{r}
install.packages("reader")
```

```{r}

#here is the link that I want to upload 
#https://raw.githubusercontent.com/wellsdata/CompText_Jour/main/data/blackindex_master.csv
```

```{r}
data <- read_csv("https://raw.githubusercontent.com/wellsdata/CompText_Jour/main/data/blackindex_master.csv")
```


```{r}
#Using code, describe the number of rows and columns in the dataset

nrow(data)
ncol(data)

# in this data set there are 1803 rows and 30 columns.. 
```


```{r}
#Create a table that displays a count of the top 5 newspaper_city entries
top_5 <- data %>%
count (newspaper_city) %>%
 arrange(desc(newspaper_city)) %>%
  slice_max(n, n=5)

# View the result
print(top_5)

# Below are the top 5 newspaper_city entries. 
```


```{r}
#Create a table that counts the different entries in the "newspaper_city" column

different_entries <- data %>%
count (newspaper_city, sort = TRUE, name = "Count") %>%
  arrange(desc(Count))

# View the result
print(different_entries)

#Below are the table that is presenting the count of the different entries "newspaper_city" column
```

```{r}
#Create a table that counts all newspaper_city entries by year. UPDATE: Filter out the NA values

entries_year <- data %>%
  filter(!is.na(newspaper_city)) %>%  # Filter out NA values
  count(newspaper_city, year, sort = TRUE, name = "Count") %>%
  arrange(desc(Count))

# View the result
print(entries_year)

#below table is providing the list of 232 entries in the form of all newspaper_city entries by year where the highest are 27 in Atlanta in 1946 and lowest is 1 in St_Louis	in 1920. 
```


```{r}
#Create a simple column ggplot chart that shows the total entries by year

  ggplot(entries_year, aes(x=year, y=Count, fill=Count))+
  geom_col(position = "dodge")+
labs(title = "TOTAL ENTRIES BY YEAR",
       subtitle = "DATA FROM 1850 TO 2002",
       caption = "Graphic by Taufiq Ahmad")

#below is the chat of total entries by year
```

#Armed with your data analysis, write a 250 word memo at the end of the R markdown document that summarizes your findings. Use this opportunity to ask questions or note any major flaws in the data.

This analysis began by loading the necessary libraries: Janitor, Tidyverse, and Rio. The dataset was initially intended to be uploaded via a provided URL link, but due to technical issues, I directly accessed the raw file from GitHub, which resolved the problem. The dataset contains 1,803 rows and 30 columns, focusing on newspaper entries categorized by city and year.

The first step was to identify the top five cities with the most newspaper entries. A major observation was the high occurrence of missing values in the "newspaper_city" column, which accounted for 1,089 entries. This presents a significant data flaw, as more than half of the entries lack city information. To extract the top cities, I utilized the slice() function, which efficiently filtered the data and revealed that Pittsburgh had the highest valid count (151 entries) after accounting for missing data.

Next, I counted the different entries in the "newspaper_city" column, resulting in 13 distinct cities. This step confirmed Pittsburgh as the city with the most entries, highlighting its prominence in the dataset once missing data was excluded.

I filtered out the missing values using the command filter(!is.na(newspaper_city)) and examined the newspaper entries by year. The analysis showed that 232 entries were spread across different years, with the highest count being 27 in Atlanta in 1946 and the lowest being a single entry in St. Louis in 1920.

A ggplot chart was developed to visualize the total entries by year. The chart revealed notable gaps in data between 1890-1904 and from 1960-2002, with a peak period of entries between 1920 and 1950, particularly in 1946.

The analysis would be enhanced by investigating the reasons behind the high newspaper activity in specific cities like Pittsburgh and Atlanta, and by examining historical events that may have influenced these trends. Additionally, addressing the extensive missing data in the "newspaper_city" column is crucial for a more accurate and comprehensive analysis.






