---
title: "NRC Sentiment_Analysis"
author: "Rob Wells"
date: "2024-8-25"
output: html_document
---

# Jour389/689 Fall 2024

Follow this script and use this dataset instead:
https://github.com/wellsdata/CompText_Jour/raw/main/data/black_press_extracted_text_june_22_2024.csv

load the following libraries: tidyverse, textdata, tidytext, quanteda, rio

```{r}



```


#import "articles_oct_19.csv" as a dataframe

```{r}
#import df created from sequence below

```

#Tokenize sentence into a df, remove stopwords


```{r}
```


# Count the words in descending order
```{r}
# Word Count

text_word_ct <- text_tokenized %>%
  count(word, sort=TRUE)
```

# NRC Sentiment

NRC Lexicon on Whole Corpus
"The nrc lexicon categorizes words in a binary fashion (“yes”/“no”) into categories of positive, negative, anger, anticipation, disgust, fear, joy, sadness, surprise, and trust."
```{r}
# cite this lexicon
#install.packages("textdata")
nrc_sentiments <- get_sentiments("nrc")
afinn_sentiments <- get_sentiments("afinn")
```

#Count the NRC sentiments
```{r}

```

### Join the NRC Sentiments with the tokenized data


```{r}

sentiments_all <- text_tokenized %>%
  inner_join(nrc_sentiments) 

```

### Count Overall Sentiment with NRC

```{r}

```

## Use ggplot to chart Sentiments with the tokenized data

```{r}

```



# Create a new dataframe just with the NRC "anger" sentiment
```{r}


```

