---
title: "Split PDFs and text"
author: "Rob Wells"
date: "2024-10-13"
output: html_document
---
Homework: Scrape PDF, compile dataframe 

Here you will work on the first part of the data processing pipeline: scraping a PDF for raw text and compiling it into a dataframe.

1) Download this PDF: https://github.com/wellsdata/CompText_Jour/blob/main/exercises/assets/pdfs/AI_yao_taufiq.PDF

2) Create a new R Markdown document.

3) Extract the text using the pdftools package

4) Split the text so you have one article per file

5) Construct a dataframe with an index of the articles a unique file name for each article

6) Pull the text articles together into a single dataframe, one row per sentence

7) BONUS NERD ZONE: Link the dataframe with the text to the index

All of the tools you need are in "Week7_split_pdfs.Rmd" and "Week8_Text_Compiler_Article_Filter_10_29.Rmd"

Please submit a link to the GitHub URL of your .rmd notebook in Elms.


```{r}
#install.packages("pdftools")
library(tidyverse)
library(pdftools)
```

# Convert PDF to text

Using 10 articles downloaded from NexisUni for Kemi Busari's research
```{r}
#Using pdftools package. Good for basic PDF extraction
#Oct 17: removed split_file folder in a cleanup
text <- pdf_text("../exercises/assets/pdfs/AI_yao_taufiq.PDF")
#pdf_text reads the text from a PDF file.
writeLines(text, "assets/extracted_text/AI_yao_taufiq.txt")
#writeLines writes this text to a text file
```


# Split text to separate articles on common identifier

In this case, NexisUni makes life easy for us. At the end of each document, there are the words "End of Document". Convenient! We search for "End of Document" and then instruct R to split the file and dump it into a standalone text file.
```{r}
# Step 1: Read the entire text file into R
#You will need to alter this for your computer
#For Mac: In Finder, Cntl + click on the filename, NOW hold down Alt/Option, and an item to copy file path will appear as Copy "Filename" as Pathname 
#https://stackoverflow.com/questions/52695546/how-to-copy-path-of-a-file-in-mac-os


file_path <- "../exercises/assets/extracted_text/AI_yao_taufiq.txt"
text_data <- readLines(file_path)

# Step 2: Combine lines into one single string
text_combined <- paste(text_data, collapse = "\n")

# Step 3: Split the text by the "End of Document" phrase
documents <- strsplit(text_combined, "End of Document")[[1]]

# Step 4: Write each section to a new file
output_dir <- "../exercises/assets/extracted_text/"
for (i in seq_along(documents)) {
  output_file <- file.path(output_dir, paste0("AI_yao_extracted_", i, ".txt"))
  writeLines(documents[[i]], output_file)
}

cat("Files created:", length(documents), "\n")
```

# Create an index from the first extracted page
--We are just grabbing the index of the 10 listed items and creating a dataframe
```{r}
yao_index <- read_lines("../exercises/assets/extracted_text/AI_yao_extracted_1.txt")
# Extract lines 16 to 58
extracted_lines <- yao_index[16:540]


# Print the extracted lines to the console
cat(extracted_lines, sep = "\n")

extracted_lines <- extracted_lines |> 
  as.data.frame() 


extracted_lines <- extracted_lines |> 
  mutate(extracted_lines = str_remove(extracted_lines, "\\| About LexisNexis \\| Privacy Policy \\| Terms & Conditions \\| Copyright © 2020 LexisNexis"))
```

#Build a final dataframe index

#?????Fix date extraction?????

```{r}
# Step 1: Trim spaces and detect rows with titles and dates
cleaned_data <- extracted_lines |>
  mutate(
    # Trim leading and trailing spaces before detection
    trimmed_line = str_trim(extracted_lines),  

    # Detect titles (start with a number and a period)
    is_title = str_detect(trimmed_line, "^\\d+\\. "),  

    # Detect dates (e.g., "Aug 14, 2024")
    is_date = str_detect(trimmed_line, "\\b\\w{3} \\d{1,2}, \\d{4}\\b")
  )

# Step 2: Shift dates to align with corresponding titles
aligned_data <- cleaned_data |>
  mutate(
    date = ifelse(lead(is_date, 1), lead(trimmed_line, 1), NA_character_)  # Shift date to title's row
  ) |>
  filter(is_title) |>
  select(trimmed_line, date)  # Keep only the relevant columns

# Step 3: Rename columns for clarity
final_data <- aligned_data |>
  rename(
    title = trimmed_line,
    date = date
  )

#Step 4: Date and Publication in separate columns, and formatted
final_data <- separate(data = final_data, col = date, into = c("date2", "publication"), sep = "  ", extra = "merge", fill = "right")


#Step 5: Format date, clean headline
final_data <- final_data |> 
  mutate(date = as.Date(date2,format = "%b %d, %Y")) |> 
  mutate(title =str_remove(title, "^\\d+\\. ")) |> 
  subset(select = -(date2)) |> 
  mutate(index = row_number()) |> 
  select(index, date, title, publication)

write_csv(final_data, "assets/yao_taufiq_index.csv")
  
```


# Raw text compiler 

###Fix Article 1 to remove the index
```{r}
article1 <- read_lines("../exercises/assets/extracted_text/AI_yao_extracted_1.txt")
# Extract lines 16 to 58
article1_cleaned <- article1[546:661]

writeLines(article1_cleaned, "../exercises/assets/extracted_text/AI_yao_extracted_1.txt")

#I copied the original extracted_1.txt and the full file and move it to another folder

```



```{r include=FALSE}
#This creates an index with the file path to the stories. And then it compiles the stories into a dataframe
#####################
# Begin SM Code #####
#####################

###
# List out text files that match pattern .txt, create DF
###

files <- list.files("../exercises/assets/extracted_text", pattern="*.txt") %>% 
  as.data.frame() |> 
  rename(filename = 1) |> 
  #create an index with the file name
 mutate(index = str_extract(filename, "\\d+")) |> 
  mutate(index = as.numeric(index))

#the actual path: #~/Code/CompText_Jour/exercises/assets/extracted_text

#Join the file list to the index

#load final data if you haven't already
#final_data <- read.csv("assets/final_data.csv")

final_index <- final_data |> 
  inner_join(files, c("index")) |> 
#you need the actual hard-coded path on this line below to the text
  mutate(filepath = paste0("/Users/robwells/Code/CompText_Jour/exercises/assets/extracted_text/", filename))

head(final_index)
```

#Text compiler
```{r}
###
# Define function to loop through each text file 
###

create_article_text <- function(row_value) {
  
  #row_value is the single argument that is passed to the function
  # Take each row of the dataframe
  temp <- final_index %>%
    slice(row_value)
  
  # Store the filename for  use in constructing articles dataframe
  temp_filename <- temp$filename
  
  # Create a dataframe by reading in lines of a given textfile
  # Add a filename column 
  articles_df_temp <- read_lines(temp$filepath) %>%
    as_tibble() %>%
    mutate(filename = temp_filename)
  
  # Bind results to master articles_df
  # <<- returns to global environment
  articles_df <<- articles_df %>%
    bind_rows(articles_df_temp)
}

###
# Create elements needed to run function
###

# Create empty tibble to store results
articles_df <- tibble()
#running once to test
#create_article_text(2) 
# Create an array of numbers to loop through, from 1 to the number of rows in our index dataframe 
row_values <- 1:nrow(final_index)

###
# Execute function using lapply
# This loops through each row of the dataframe and append results to master file
###

lapply(row_values, create_article_text)

###
# Clean up articles_df and join to index dataframe
###

articles_df <- articles_df %>%
  select(filename, sentence=value) %>%
  inner_join(final_index)

#After viewing articles_df, I see 64 lines from the index that I don't need. Cutting them 

articles_df <- articles_df %>%
  slice(-c(1:64)) |> 
  #gets rid of blank rows
    filter(trimws(sentence) != "") 

write.csv(articles_df, "../exercises/assets/extracted_text/yao_taufiq_df.csv")

```