---
title: "Case Study case_when: Milwaukee Evictions"
author: "Rob Wells"
date: "11/17/2024"
output: html_document
---

# Analyzing Eviction Court Data From Milwaukee
### Data analysis by Rob Wells


- **The records were scraped from this website**  
https://wcca.wicourts.gov/advanced.html 

--------------------------------------------------------------------
- **Part 1: Load Libraries and Data**

```{r include=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
library(kableExtra)
```
Load Milwaukee City and County Data for Evictions
```{r}
milw <- rio::import("../data/sta_wi_milwaukee_2019_2020.csv") 
```

- **Clean names**
```{r}
milw <- janitor::clean_names(milw)
```

```{r}
glimpse(milw)
```

- **Parse Cities **
```{r}
milw <- milw |> 
  mutate(defendant_address2 = defendant_address) |> # makes a copy of a column you will now pick apart
  separate(defendant_address, into = c("address", "city", "state_zip"), sep = ", ", extra = "merge", fill = "right")

```


```{r}
milw %>% 
  group_by(city) %>%
  count(city) %>% 
  arrange(desc(n))
```

- **Process dates** 
```{r}
milw <- milw |> 
  mutate(date = ymd(file_date)) |> 
  mutate(yearmon = format(date, "%Y-%m"))
#Filter to July 31, 2000
milw <- milw %>%
filter(file_date <= as.Date("2020-07-31")) %>%
distinct()
```

### Total Eviction Filings by Month in Milwaukee City and County, Jan. 1 2019-July 31, 2020
**Wisconsin eviction moratorium was March 27 - May 27, 2020**
```{r}
z <- milw %>% 
  group_by(yearmon) %>%
  count(yearmon)
z
```

- **Context: Total eviction filings in Wisconsin City and County since Jan. 1, 2019: 19,420**
```{r}
summary(z$n)
sum(z$n)
```

- **Chart of Milwaukee City and County Eviction Filings, Jan 1. 2019-July 31, 2020**  
**Wisconsin eviction moratorium was March 27 - May 27, 2020**
```{r}

ggplot(z, aes(x=yearmon, y=n, fill=yearmon)) +
  geom_col() +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title="Milwaukee City-County Eviction Filings Per Month Jan. 2019-July 2020",y="Evictions", x= "", 
       caption = "Graphic by Rob Wells. 8-21-2020", subtitle = "Source: Wisconsin Circuit Court records")+
  theme(legend.position="none") 
#ggsave("Milw_Evict_8_21_2020.png",device = "png",width=9,height=6, dpi=500)
```

- **Clean the columns - remove stray spaces, numerals**

```{r}
milw <- milw |> 
  mutate(
    plaintiff_one = str_trim(plaintiff) |> 
      str_remove_all("[[:punct:]]") |> 
      str_remove_all("[0-9]"),
    defendant_one = str_trim(defendant) |> 
      str_remove_all("[[:punct:]]")
  )
```


- **Summary of Milwaukee Plaintiffs **
```{r}
#Build a simple summary table
plaintiff1 <- milw %>% 
  select(plaintiff_one, defendant_one, date) %>% 
         group_by(plaintiff_one) %>% 
         count(plaintiff_one) %>% 
         ungroup() %>% 
         arrange(desc(n))

#write.csv(plaintiff1, "milplaintiff.7.6.2020.csv")


plaintiff1
```
##Cleaning of Plaintiffs Names, Normalizing and Grouping
```{r}
milw <- milw %>% 
  mutate(plaintiff_fixed = case_when(plaintiff_one=="meaux park llc" ~ "berrada",
                                     plaintiff_one=="berrada properties  llc" ~ "berrada",
                                     plaintiff_one=="spring court apartments llc" ~ "berrada",
                                     plaintiff_one=="dineen park townhomes llc" ~ "berrada",
                                     plaintiff_one=="arbor gardens apartments llc" ~ "berrada",
                                          TRUE~plaintiff_one)) 
```

Tally
```{r}
milw |> 
  count(plaintiff_fixed) |> 
  arrange(desc(n))


```



# Your turn
Categorize the following as "berrada" properties
"luther haven llc" 
"willow court apartments llc" 
"garden terrace homes llc" 
"good hope townhomes ii llc" 
"good hope townhomes llc" 
"colonial pointe ii apartments llc" 
"colonial pointe il apartments llc" 
"colonial pointe i apartments llc" 

Categorize the following as  "anchor_properties"
"atlantis properties llc" 
"anchor properties llc" 
 "action properties llc" 
"afc properties llc" 
"atunna properties llc" 
"affordable properties llc" 
"acme properties  llc" 


```{r}
#Save as a csv
#write.csv(p2, "milw_landlord_cleaned_8_6_2020.csv")
```


- **Context: Berrada was responsible for 17% of eviction filings in Milwaukee City and County since March 27, 2020.** Since the moratorium started in March 27, 2020, there have been 2,929 evictions. Berrada and its 18 affiliated entities had 484 eviction filings, or 17% of the total. Anchor and its affiliates were responsible for 4% of eviction filings.
```{r}
#Cleaned Milwaukee Plaintiffs since March 27 2020
top2 <- p2 %>% 
  filter(date > '2020-03-27') %>% 
  select(plaintiff_fixed, yearmon) %>% 
         group_by(plaintiff_fixed) %>% 
         count(plaintiff_fixed) %>% 
         filter(n > 5) %>% 
         ungroup() %>% 
         arrange(desc(n))
top2
```


- **Top with top 50 Results of Milwaukee Landlords**
```{r}
top2 %>% 
  top_n(50, n) %>% 
  kable() %>% 
  kable_styling("striped")

```




**--30--**

Don't Peek!


- **Rename and Group Landlords, Substitute top names and normalize**
```{r}
p2 <- p2 %>% 
  mutate(plaintiff_fixed = case_when(plaintiff_one=="meaux park llc" ~ "berrada",
                                     plaintiff_one=="berrada properties  llc" ~ "berrada",
                                     plaintiff_one=="spring court apartments llc" ~ "berrada",
                                     plaintiff_one=="dineen park townhomes llc" ~ "berrada",
                                     plaintiff_one=="arbor gardens apartments llc" ~ "berrada",
                                     plaintiff_one=="luther haven llc" ~ "berrada",
                                     plaintiff_one=="willow court apartments llc" ~ "berrada",
                                     plaintiff_one=="garden terrace homes llc" ~ "berrada",
                                     plaintiff_one=="good hope townhomes ii llc" ~ "berrada",
                                     plaintiff_one=="good hope townhomes llc" ~ "berrada",
                                     plaintiff_one=="colonial pointe ii apartments llc" ~ "berrada",
                                     plaintiff_one=="colonial pointe il apartments llc" ~ "berrada",
                                     plaintiff_one=="colonial pointe i apartments llc" ~ "berrada",
                                     plaintiff_one=="mariners pointe apartments llc" ~ "berrada",
                                     plaintiff_one=="berrada properties llc" ~ "berrada",
                                     plaintiff_one=="yb florist llc" ~ "berrada",
                                     plaintiff_one=="yb medford llc" ~ "berrada",
                                     plaintiff_one=="yb silver spring llc" ~ "berrada",
                                     plaintiff_one=="yb victory llc" ~ "berrada",
                                     plaintiff_one=="falcon crest llc" ~ "berrada",
                                     plaintiff_one=="wisniewski michael s" ~ "michael_s_wisniewski",
                                     plaintiff_one=="bruckner daniel w" ~ "daniel_w_bruckner",
                                     plaintiff_one=="nawrocki gregory w" ~ "gregory_w_nawrocki",
                                     plaintiff_one=="nawrocki gregory" ~ "gregory_w_nawrocki",
                                     plaintiff_one=="nawrocki gragory w" ~ "gregory_w_nawrocki",
                                     plaintiff_one=="martin bruce" ~ "bruce_martin",
                                     plaintiff_one=="housing authority of the city of milwaukee" ~ "milwaukee_city_housing_authority",
                                     plaintiff_one=="city of milwaukee" ~ "milwaukee_city",
                                     plaintiff_one=="city  of milwaukee" ~ "milwaukee_city",
                                     plaintiff_one=="milwaukee living llc" ~ "milwaukee_living",
                                     plaintiff_one=="milwaukee capital properties llc" ~ "milwaukee_capital_properties",
                                     plaintiff_one=="good hope townhomes ii llc" ~ "good_hope_townhomes",
                                     plaintiff_one=="good hope townhomes llc" ~ "good_hope_townhomes",
                                     plaintiff_one=="howard and th street holding company" ~ "howard_76_st_holding_co",
                                     plaintiff_one=="maas robert" ~ "robert_maas",
                                     plaintiff_one=="maas robert j" ~ "robert_maas",
                                     plaintiff_one=="atlantis properties llc" ~ "anchor_properties",
                                     plaintiff_one=="anchor properties llc" ~ "anchor_properties",
                                      plaintiff_one=="action properties llc" ~ "anchor_properties",
                                     plaintiff_one=="afc properties llc" ~ "anchor_properties",
                                     plaintiff_one=="atunna properties llc" ~ "anchor_properties",
                                     plaintiff_one=="affordable properties llc" ~ "anchor_properties",
                                     plaintiff_one=="acme properties  llc" ~ "anchor_properties",
                                     plaintiff_one=="alotta properties llc" ~ "anchor_properties",
                                     plaintiff_one=="anumbera properties llc" ~ "anchor_properties",
                                     plaintiff_one=="allsortsa properties llc" ~ "anchor_properties",
                                     plaintiff_one=="absolute properties llc" ~ "anchor_properties",
                                    TRUE~plaintiff_one)) 

```
