---
title: "r-load-analyze-exercise"
author: "Rob Wells"
date: "2024-08-23"
output: html_document
---
---
editor: 
  markdown: 
    wrap: 72
---

# Loading and Analyzing Data Exercise

This class exercise is a companion to the following chapter in Computational Text Analysis: [Ch. 11  Loading and Analyzing Data](https://wellsdata.github.io/CompText_Jour/docs/r-load-analyze-visualize.html)


#1: load tidyverse, rio and janitor libraries
```{r}


```


#2: Import Data:

https://github.com/wellsdata/CompText_Jour/raw/main/data/mob_analysis_data.csv
```{r}


```

How many rows?

```{r}

```

How many columns?

```{r}

```

Dimensions: Gives number rows, then columns

```{r}
```

Names of your columns

```{r}
```

Check data types

```{r}

```

Let's look at the first six rows

```{r}

```

Use summary to view the range of your data

```{r}

```

Create a table that displays a count of the top 15 newspaper entries

```{r}

```

Create a table that counts the different entries in the "keyword" column

```{r}

```

Create a table that counts all entries by year

```{r}

```

Create a simple column ggplot chart that shows the total entries by year

```{r}

```

