---
title: "homework-basic-r-proficiency"
author: "Rob Wells"
date: "2024-08-23"
output: html_document
---

# Homework - Basic R Proficiency

This class exercise is a companion to the following chapter in Computational Text Analysis: [Ch. 11  Loading and Analyzing Data](https://wellsdata.github.io/CompText_Jour/docs/r-load-analyze-visualize.html)


#1: load tidyverse, rio and janitor libraries
```{r}


```


#2: Import Data using read.csv:

https://github.com/wellsdata/CompText_Jour/raw/main/data/mob_analysis_data.csv
```{r}

```

#3: Provide dimensions: Gives number rows, then columns

```{r}
```

#4: Show names of your columns

```{r}

```

#5: Check data types

```{r}

```

#6: Display the first six rows

```{r}

```


#7: Create a table counting the number of entries by year and display the top 5 results

```{r}


```

#8: Create a table that counts the different entries in the "keyword" column

```{r}


```


#9: Create a simple column ggplot chart that shows the total entries by year

```{r}

```

