---
title: "homework-basic-r-proficiency"
author: "Rob Wells"
date: "2024-08-23"
output: html_document
---


# Homework - Basic R Proficiency

This class exercise is a companion to the following chapter in Computational Text Analysis: [Ch. 11  Loading and Analyzing Data](https://wellsdata.github.io/CompText_Jour/docs/r-load-analyze-visualize.html)


#1: load tidyverse, rio and janitor libraries
```{r}
library(tidyverse)
library(janitor)
library(rio)
```

```{r}
install.packages("tidyverse")
install.packages("janitor")
install.packages("rio")
```

#2: Import Data using read.csv:

https://github.com/wellsdata/CompText_Jour/raw/main/data/mob_analysis_data.csv
```{r}
Data <- read.csv("https://github.com/wellsdata/CompText_Jour/raw/main/data/mob_analysis_data.csv")
```

```{r}
head(Data)
```

```{r}
summary(Data)
```

```{r}
str(Data)
```

#3: Provide dimensions: Gives number rows, then columns

```{r}
dim(Data)
nrow(Data)
ncol(Data)
```

```{r}
# dimensions of the dataset
dimensions <- dim(Data)
print(dimensions)  

# number of rows
num_rows <- nrow(Data)
print(num_rows) 

# number of columns
num_columns <- ncol(Data)
print(num_columns)
```

#4: Show names of your columns

```{r}
colnames(Data)
```

#5: Check data types

```{r}
str(Data)
```


#6: Display the first six rows

```{r}
head(Data)
```


#7: Create a table counting the number of entries by year and display the top 5 results

```{r}
head(Data$Year)
```

```{r}

# Check for any conversion issues
summary(Data$Year)

```

```{r}
format(Data$Year)
```
```{r}
table(Data$Year)
```

```{r}
#number of entries by year
year_counts <- Data %>%
  count(Year) %>%
  arrange(desc(n)) %>%
  head(5)

# View the result
print(year_counts)
```

#8: Create a table that counts the different entries in the "keyword" column

```{r}
colnames(Data)
```

```{r}
# Different entries in the 'keyword' column
keyword_counts <- Data %>%
  count(keyword) %>%
  arrange(desc(n))

# View the result
print(keyword_counts)
```


#9: Create a simple column ggplot chart that shows the total entries by year

```{r}
#number of entries by year
Total_year_counts <- Data %>%
  count(Year) %>%
  arrange(desc(n)) %>%

# View the result
print(Total_year_counts)
```

```{r}
# Count the number of entries by year
 Data %>%
  count(Year) %>%
  arrange(desc(n))

# Create and display the column chart
ggplot(Total_year_counts, aes(x = Year, y = n)) +
  geom_col(fill = "steelblue") +
  labs(title = "Total Entries by Year", x = "Year", y = "Total Entries") +
  theme_minimal()
```


