---
title: "Week 4 Bigrams Homework"
author: "Rob Wells"
date: '2025-02-20'
output: html_document
---

# Jour 689 Spring 2025:

In this exercise, you will import the Nexis spreadsheet and create bigrams from the headlines

Setup. These instructions are important. Follow them carefully

1) Create a new folder called "bigrams_week_5" in your class folder
2) Copy this file to "bigrams_week_5"
3) Copy your spreadsheet to "bigrams_week_5"
4) Create an .Rproj file that points to "bigrams_week_5"

```{r}
#load tidyverse, tidytext, rio, janitor libraries


```

```{r}
#Import spreadsheet using rio::import and clean the names


XXXX <- rio::import("XXXX.XLSX") 

```
# Tokenize the hlead column

Copy the code from the in-class bigrams exercise and tokenize just the hlead column from your dataframe

Hint: you're changing just one variable

```{r}

```

#Remove stopwords and count the words
```{r}

data(stop_words)



# Word Count


```

# Create Bigrams

```{r}


#Filter out stop words.





```


# Write 250 words about what your learned analyzing the list of tokens and bigrams. Include questions about the process and any difficulties you encountered.

