---
title: "Bigrams Exercise Sept 24"
author: "Bridget Lang"
date: '2024-09-28'
output: html_document
---

# Jour 389/689 Fall 2024:


```{r}
#load tidyverse, tidytext, rio and quanteda libraries
library(tidyverse)
library(tidytext)

```

```{r}
#Import dataframe 

lynch <- read_csv("../data/articles_oct_19.csv")

```


# Create a new dataframe that filters articles for 1900 to 1910

```{r}

firstDecade <- lynch %>%
  filter(year >= 1900) %>%
  filter(year <= 1910)
```


# Count the number of distinct articles in 1900 dataframe
```{r}
firstDecade %>% 
  select(filename) %>% 
  distinct(filename, .keep_all = TRUE) %>% 
  count(filename) %>% 
  summarize(total =sum(n)) 
#1732 distinct articles
```

# Count the number of newspaper_states in the 1900 corpus
```{r}
firstDecade %>% 
  select(newspaper_state, filename) %>% 
  distinct(filename, .keep_all = TRUE) %>% 
  count(newspaper_state) %>% 
  arrange(desc(n))


```

# Tokenize the 1900 lynching stories

```{r}
stories <- str_replace_all(firstDecade$sentence, "- ", "")
stories_df <- tibble(stories,)

stories_tokenized <- stories_df %>%
  unnest_tokens(word,stories)

stories_tokenized

```


#Remove stopwords
The tidytext package includes the stop_words dataset.It contains, as of this writing, 1,149 words that data scientists and linguistic nerds felt could be removed from sentences because they don't add meaning. Filtering out these words can help focus on the more meaningful content, making it easier to uncover trends, themes, and key information in large amounts of text. Obviously, we have different priorities and we may or may not want to use stop_words or we have want to provide a customized list of stop words.

The stop_words list is derived from three separate lists, or lexicons: SMART (571 words), onix (404 words), and snowball (174 words)

The ONIX lexicon comes from the Open Information Exchange and is often used in text mining and natural language processing. 

The Snowball lexicon is part of a broader project that has algorithms that simplify words in different languages by reducing them to their root form. It's best known for the Porter stemming algorithm, which, for example, changes "running" to "run." 

Lastly, the SMART lexicon is a set of common words, like "and," "the," and "is," and it comes from the SMART Information Retrieval System, created at Cornell University in the 1960s.

```{r}
data(stop_words)

test <- stop_words %>% 
  as.data.frame()

head(test)
```
# Strip out stop words

```{r}

stories_tokenized <- stories_tokenized %>%
  anti_join(stop_words, by = c("word" = "word")) %>%
  filter(word != "temp_file") %>%
  filter(word != "stories_corpus") %>%
  filter(!grepl('[0-9]', word))

# Word Count

story_word_ct <- stories_tokenized %>%
  count(word, sort=TRUE)

head(story_word_ct)


```

# Bigrams
## We are now creating two word phrases but before the stop words are taken out

```{r}
stories_bigrams <- stories_df %>%
  unnest_tokens(bigram, stories, token="ngrams", n=2)

stories_bigrams_separated <- stories_bigrams %>%
  separate(bigram, c("word1", "word2"), sep = " ")
```

# Create a new dataframe with counts of the bigrams
```{r}

stories_bigrams_count <- stories_bigrams %>%
  count(bigram, sort=TRUE)

```

## Now filter the counts 
```{r}

stories_bigrams_filtered <- stories_bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)

stories_bigram_cts2 <- stories_bigrams_filtered %>%
  count(word1, word2, sort = TRUE) %>% 
  filter(!is.na(word1))

stories_bigram_cts2
```

# Add a "1900" decade column

Hint: use mutate

```{r}
library(dplyr)
lynch <- lynch %>%
  mutate(firstDecade = ifelse(year >= 1900 & year <= 1910, 'Y', "N"))

```


# YOUR TURN

Create one dataframe with black press articles
Create a second dataframe without black press articles
Produce the top 20 bigrams for the black press and non-black press coverage
Compare and discuss!

```{r}
#black press
black_press <- lynch %>%
  filter(!is.na(black_press))
```

```{r}
#tokenize
black_press_stories <- str_replace_all(black_press$sentence, "- ", "")
black_press_stories_df <- tibble(black_press_stories,)

black_press_stories_tokenized <- black_press_stories_df %>%
  unnest_tokens(word,black_press_stories)

black_press_stories_tokenized

```

```{r}
#strip out stop words
data(stop_words)

black_press_stories_tokenized <- black_press_stories_tokenized %>%
  anti_join(stop_words, by = c("word" = "word")) %>%
  filter(word != "temp_file") %>%
  filter(word != "stories_corpus") %>%
  filter(!grepl('[0-9]', word))

# Word Count

black_press_story_word_ct <- black_press_stories_tokenized %>%
  count(word, sort=TRUE)

```

```{r}
#bigrams before stripping stop words

black_press_stories_bigrams <- black_press_stories_df %>%
  unnest_tokens(bigram, black_press_stories, token="ngrams", n=2)

black_press_stories_bigrams_separated <- black_press_stories_bigrams %>%
  separate(bigram, c("word1", "word2"), sep = " ")

```

```{r}
black_press_stories_bigrams_count <- black_press_stories_bigrams %>%
  count(bigram, sort=TRUE)
```

```{r}
#bigrams with stpo worfs filtered
black_press_stories_bigrams_filtered <- black_press_stories_bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)

black_press_stories_bigram_cts2 <- black_press_stories_bigrams_filtered %>%
  count(word1, word2, sort = TRUE) %>% 
  filter(!is.na(word1))

```

```{r}
#not black press
not_black_press <- lynch %>%
  filter(is.na(black_press))
```

```{r}
#tokenize
not_black_press_stories <- str_replace_all(not_black_press$sentence, "- ", "")
not_black_press_stories_df <- tibble(not_black_press_stories,)

not_black_press_stories_tokenized <- not_black_press_stories_df %>%
  unnest_tokens(word, not_black_press_stories)

```

```{r}
#strip out stop words

not_black_press_stories_tokenized <- not_black_press_stories_tokenized %>%
  anti_join(stop_words, by = c("word" = "word")) %>%
  filter(word != "temp_file") %>%
  filter(word != "stories_corpus") %>%
  filter(!grepl('[0-9]', word))

# Word Count

not_black_press_story_word_ct <- not_black_press_stories_tokenized %>%
  count(word, sort=TRUE)

```

```{r}
#bigrams before stripping stop words

not_black_press_stories_bigrams <- not_black_press_stories_df %>%
  unnest_tokens(bigram, not_black_press_stories, token="ngrams", n=2)

not_black_press_stories_bigrams_separated <- not_black_press_stories_bigrams %>%
  separate(bigram, c("word1", "word2"), sep = " ")

```

```{r}
not_black_press_stories_bigrams_count <- not_black_press_stories_bigrams %>%
  count(bigram, sort=TRUE)
```

```{r}
#bigrams with stop words filtered
not_black_press_stories_bigrams_filtered <- not_black_press_stories_bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)

not_black_press_stories_bigram_cts2 <- not_black_press_stories_bigrams_filtered %>%
  count(word1, word2, sort = TRUE) %>% 
  filter(!is.na(word1))

```
# Top 20 bigrams for black press 
```{r}
head(black_press_stories_bigram_cts2, 20)
```
# Top 20 bigrams for non-black press 
```{r}
#top 20 bigrams for non black press 
head(not_black_press_stories_bigram_cts2, 20)
```

The black press seem to cover more about people of color in general, with there obviously being a higher number of articles about any of these topics. The most used bigram by the black press is "negro lynched" which indicates a very high coverage of individual lynchings or crimes again black people. The second most common bigram is "county jail" which indicated a high amount of coverage of black people being incarcerated or accused of crimes. The third most common is "white woman", which is interesting given the role that white women and their accusation have played in the historical (and current) persecution of black men in particular. 

The non-black press's most used bigram was "colored people", followed by "lynch law", then "national association". The terms more often used by the non-black press seem as though they would be used in a more general context, as if describing a group of people or a common issue rather than reporting on specific incidents. The bigram "national association" is likely the beginning of the NAACP. According to blackpast.org, "lynch law" describes a set of beliefs or unwritten rules that people believed justified the use of violence against black people, as they felt the government did not do enough (https://www.blackpast.org/african-american-history/1900-ida-b-wells-lynch-law-america/) The fact that this is a primarily used phrase in the non-black press could lead one to make the conclusion that the non-black press are justifying the act or use of lynching. However, it is imporant to remember that it is simply a quantitative piece of data, and looking into the context of the use of this specific bigram is important before drawing any specific conlusions from this information 
