---
title: "NRC Sentiment_Analysis"
author: "Rob Wells"
date: "2024-8-25"
output: html_document
---

# Jour389/689 Fall 2024

Follow this script and use this dataset instead: <https://github.com/wellsdata/CompText_Jour/raw/main/data/black_press_extracted_text_june_22_2024.csv>

load the following libraries: tidyverse, textdata, tidytext, quanteda, rio

```{r}
library(tidyverse)
library(textdata)
library(tidytext)
library(quanteda)
library(rio)

```

# Tokenizing data

```{r}
#import df created from sequence below

lynch <- read.csv("../data/articles_oct_19.csv")


#update to the file path for your code folder
#black <- read.csv("~/Code/CompText_Jour/data/black_press_extracted_text_june_22_2024.csv")
```

#Tokenize sentence into a df, remove stopwords

```{r}
text_tokenized1 <- lynch %>% 
  select(sentence) %>% 
  unnest_tokens(word, sentence) 
```


```{r}
text_tokenized <- lynch %>% 
  select(sentence) %>% 
  mutate(sentence = str_replace_all(sentence, "- ", "")) %>% 
  unnest_tokens(word, sentence) %>% 
  filter(!word %in% stop_words$word) %>% 
  filter(word != "stories_corpus") %>%
  filter(!grepl('[0-9]', word))
```

# Count the words in descending order

```{r}
text_word_ct <- text_tokenized %>%
  count(word, sort=TRUE)

text_word_ct
```

# NRC Sentiment

NRC Lexicon on Whole Corpus "The nrc lexicon categorizes words in a binary fashion (“yes”/“no”) into categories of positive, negative, anger, anticipation, disgust, fear, joy, sadness, surprise, and trust."

```{r}
# cite this lexicon
#install.packages("textdata")
nrc_sentiments <- get_sentiments("nrc")
afinn_sentiments <- get_sentiments("afinn")

nrc_sentiments %>% count(sentiment)

#sentiment & count
# anger	1246			
# anticipation	837			
# disgust	1056			
# fear	1474			
# joy	687			
# negative	3318			
# positive	2308			
# sadness	1187			
# surprise	532			
# trust	1230	

nrc_sentiments %>% 
  group_by(word) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  distinct()
```

### Review NRC Overall Sentiment

```{r}

sentiments_all <- text_tokenized %>%
  inner_join(nrc_sentiments) 

#this dictionary assigns different values to the same word. negro is negative, sadness whereas lynch is anger, disgust, fear, negative and sadness.



sentiments_all %>% 
  group_by(word) %>% 
    count(sentiment) %>% 
  arrange(desc(n))

```

### Count Overall Sentiment with NRC

```{r}
sentiments_all <- text_tokenized %>%
  inner_join(nrc_sentiments) %>%
  count(sentiment, sort = TRUE) %>% 
  mutate(pct_total =round(n/sum(n), digits=2))

sentiments_all


```

## Figure 5: Sentiment chart

```{r}
library(ggplot2)

afinn_plot <- sentiments_all %>% 
  ggplot(aes(x = sentiment, y = n,fill = n)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
  labs(title = "Total Sentiment in Black Press Lynching News Coverage",
       subtitle = " ",
       caption = "NRC Sentiment analysis. Graphic by Rob Wells, 8-25-2024",
       y="Score",
       x="total sentiment score")

afinn_plot + scico::scale_fill_scico(palette = "vik")

afinn_plot
# ggsave("Figure5_afinn_sentiment_jan2.png",device = "png",width=9,height=6, dpi=800)


```

```{r}
# Anger

nrc_anger <- nrc_sentiments %>%
  filter(sentiment == "anger")

lynching_anger <- text_tokenized %>%
  inner_join(nrc_anger) %>%
  count(word, sort = TRUE)

lynching_anger

```

```{r}
# Anticipation
# results / themes not as clear as anger

nrc_anticipation <- nrc_sentiments %>%
  filter(sentiment == "anticipation")

lynching_anticipation <-text_tokenized%>%
  inner_join(nrc_anticipation) %>%
  count(word, sort = TRUE)

lynching_anticipation

```

```{r}
# Fear
# see a reflection of the basic word count in these results

nrc_fear <- nrc_sentiments %>%
  filter(sentiment == "fear")

lynching_fear <-text_tokenized%>%
  inner_join(nrc_fear) %>%
  count(word, sort = TRUE)

lynching_fear

```

```{r}
# Disgust
# see a reflection of the basic word count in these results

nrc_disgust <- nrc_sentiments %>%
  filter(sentiment == "disgust")

lynching_disgust <-text_tokenized%>%
  inner_join(nrc_disgust) %>%
  count(word, sort = TRUE)

lynching_disgust

```

### Create custom dictionary

```{r}
#We should do our own custom sentiment dictionary based on the top 500 words. 

text_500 <- text_word_ct %>% 
  filter(n >= 29)

custom_dictionary <- text_500 %>%
  inner_join(nrc_sentiments)

#write.csv(custom_dictionary, "custom_dictionary.csv")

```

# Extra Content:

### Here are some ideas on what you can do by decade and region

# Sentiment by decade

```{r}
#before 1870
pre1870 <- lynch %>% 
  filter(year < 1870)

#1870-1879
the1870s <-  lynch %>% 
  filter(year >= 1870 & year <=1879)

#1880-1889
the1880s <-  lynch %>% 
  filter(year >= 1880 & year <=1889)

#1890-1899
the1890s <-  lynch %>% 
  filter(year >= 1890 & year <=1899)

#1900-1909
the1900s <-  lynch %>% 
  filter(year >= 1900 & year <=1909)

#1910-1919
the1910s <-  lynch %>% 
  filter(year >= 1910 & year <=1919)

#1920-1929
the1920s <-  lynch %>% 
  filter(year >= 1920 & year <=1929)

#1930-1960
post1930s <-  lynch %>% 
  filter(year >= 1930)

```

#### Compile by decade

```{r}
lynch_decade <- lynch %>% 
  mutate(decade = case_when(
         year < 1870 ~ "pre1870",
        year >= 1870 & year <=1879 ~ "1870s",
         year >= 1880 & year <=1889 ~ "1880s",
         year >= 1890 & year <=1899 ~ "1890s",
        year >= 1900 & year <=1909 ~ "1900s",
        year >= 1910 & year <=1919 ~ "1910s",
        year >= 1920 & year <=1929 ~ "1920s",
        year >= 1930 ~ "post1930s"
         ))

```

### Regional classification

```{r}
# lynch_decade %>% 
#   count(newspaper_state)
lynch_decade <- lynch_decade %>% 
  mutate(region=newspaper_state) %>% 
  mutate(region = case_when(region=="South Carolina" ~ "South",
                            region=="Texas" ~ "South",
                            region=="Louisiana" ~ "South",
                            region=="Tennessee" ~ "South",
                            region=="Mississippi" ~ "South",
                            region=="Arkansas" ~ "South",
                            region=="Alabama" ~ "South",
                            region=="Georgia" ~ "South",
                            region=="Virginia" ~ "South",
                            region=="Florida" ~ "South",
                            region=="North Carolina" ~ "South",
                            region=="Maryland" ~ "Border",
                            region=="Delaware" ~ "Border",
                            region=="West Virginia" ~ "Border",
                            region=="Kentucky" ~ "Border",
                            region=="Missouri" ~ "Border",
                            region=="Maine" ~ "North",
                            region=="New York" ~ "North",
                            region=="New Hampshire" ~ "North",
                            region=="Vermont" ~ "North",
                            region=="Massassachusetts" ~ "North",
                            region=="Connecticut" ~ "North",
                            region=="Rhode Island" ~ "North",
                            region=="Pennsylvania" ~ "North",
                            region=="New Jersey" ~ "North",
                            region=="Ohio" ~ "North",
                            region=="Indiana" ~ "North",
                            region=="Kansas" ~ "North",
                            region=="Michigan" ~ "North",
                             region=="Wisconsin" ~ "North",
                             region=="Minnesota" ~ "North",
                             region=="Iowa" ~ "North",
                             region=="California" ~ "North",
                             region=="Nevada" ~ "North",
                             region=="Oregon" ~ "North",
                            region=="Illinois" ~ "North",
                            region=="Nebraska" ~ "Misc",
                            region=="Colorado" ~ "Misc",
                            region=="North Dakota" ~ "Misc",
                            region=="South Dakota" ~ "Misc",
                            region=="Montana" ~ "Misc",
                            region=="Washington" ~ "Misc",
                            region=="Idaho" ~ "Misc",
                            region=="Wyoming" ~ "Misc",
                            region=="Utah" ~ "Misc",
                            region=="Oklahoma" ~ "Misc",
                            region=="New Mexico" ~ "Misc",
                            region=="Arizona" ~ "Misc",
                            region=="Alaska" ~ "Misc",
                            region=="Hawaii" ~ "Misc",
                            region=="District of Columbia" ~ "Misc",
                            region=="Virgin Islands" ~ "Misc",
                                                     TRUE~region)) 


```

Overall, negative sentiment was 70% of the Southern newspaper coverage but just 69% of the Northern newspaper coverage.

```{r}
sentiments_south <- sentiments_south %>% 
  rename(south_n = n, south_pct = pct_total)

sentiments_north <- sentiments_north %>% 
  rename(north_n = n, north_pct = pct_total)

sentiments_border <- sentiments_border %>% 
  rename(border_n = n, border_pct = pct_total)

sentiments_misc <- sentiments_misc %>% 
  rename(misc_n = n, misc_pct = pct_total)



sent_regions <- sentiments_south %>% 
  inner_join(sentiments_north) %>% 
  inner_join(sentiments_border) %>% 
  inner_join(sentiments_misc)

#write.csv(sent_regions, "sent_regions_jan6.csv")
```
