---
title: "Metadata"
author: "Wells"
date: "2026-09-15"
output: html_document
---

Class exercise to import your project metadata from the spreadsheet into R

--Create a folder called "class_project" in your code folder

--Save this file to the "class_project" folder

--Move your Nexis spreadsheet to the folder

--Create a project file on R Studio: File \| New Project \| \*\*\* select Existing folder \| "class_project"

--After project.Rproj launches, then launch this file, metadata.Rmd, and complete the tutorial

**By doing this, you avoid the ring of hell where R can't find your data**

#load tidyverse, janitor, rio

```{r}

library(tidyverse)
library(janitor)
library(rio)
#install.packages("readxl")
library(readxl)

```

#Import your data

```{r}

data <- read_excel("benji_data.XLSX") |> 
  clean_names()

View(data)

names(data)
```






# Clean names

```{r}
data <- data |> 
  clean_names()

```

# check data types

```{r}
glimpse(data)
```

Next Steps: Count years, and create a ggplot

# Transform date into years. See tutorial

```{r}

data <- data |> 
  mutate(year = year(published_date))

```


#count
```{r}

data |> 
  count(year)
```

#publication
```{r}

data |> 
  count(publication_4) |> 
  arrange(desc(n))
```

# counting
```{r}
data_big <- data |> 
  mutate(length1 = as.numeric(length)) |> 
  filter(length1 > 2093 & length1 <2505)

```



# Create a ggplot

```{r}
data |> 
  count(year) |> 
   ggplot(aes(x = year, y = n,
             fill = n)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
  scale_x_continuous(breaks=c(2023:2026)) + #fixes wonky x axis
  labs(title = "Benji Data",
       subtitle = "Limited NexisUni Run",
       caption = "Graphic by Rob Wells, 9-15-2026",
       y="Number of Articles",
       x="")



```

#Count publications

```{r}

```

#Visualize publications

```{r}

```

# NOTES ON FIXING COMPLEX Date COLUMN WITH HOURS AND MINUTES

```{r}
women <- women |> 
  mutate(date = as.Date(strptime(published_date, format = "%B %d, %Y %A")))
```



```{r}
nobel <- read_excel("nobel.XLSX") |> 
  clean_names() |> 
  mutate(year = year(published_date))


```


```{r}
nobel |> 
  count(publication_4) |> 
   # ggplot(aes(x = publication_4, y=n, 
   #           fill = n)) +
    ggplot(aes (x = reorder(publication_4, n), y = n, fill = n))  +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
#theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  #scale_x_continuous(breaks=c(2021:2026)) + #fixes wonky x axis
    coord_flip() +
  labs(title = "Nobel Data",
       subtitle = "Limited NexisUni Run",
       caption = "Graphic by Rob Wells, 9-15-2026",
       y="Publications",
       x="Number of Articles")



```
