---
title: "Metadata"
author: "Wells"
date: "2026-09-15"
output: html_document
---

Class exercise to import your project data into R --Create a folder called "class_project" in your code folder --Save this file to the "class_project" folder --Move your Nexis spreadsheet to the folder --Create a project file on R Studio: File \| New Project \| \*\*\* select Existing folder \| "class_project" --After project.Rproj launches, then launch this file, metadata.Rmd, and complete the tutorial

**By doing this, you avoid the ring of hell where R can't find your data**

#load tidyverse, janitor, rio

```{r}

library(tidyverse)
library(janitor)
library(rio)
```

#Import your data

```{r}

#bmore <- rio::import("")
library(readxl)
women <- read_excel("xinjiang_women.XLSX")
View(women)

```

# Clean names

```{r}
women <- women |> 
  clean_names()

```



# My claude.ai query and reply:

Q: In r, i need to create a new properly formatted date column from a "published_date" column with character data that looks like this: "January 14, 2025 Tuesday", - how do I transform it into a date format?

A: You can use the as.Date() function along with strptime() to convert that date format. Here's how:

df \<- df \|\> mutate(new_date = as.Date(strptime(published_date, format = "%B %d, %Y %A")))

The format string specifies:

%B - full month name (January) %d - day of the month (14) %Y - 4-digit year (2025) %A - full weekday name (Tuesday)

This will create a proper Date column that you can use for analysis and plotting. The resulting date will be in the standard format "2025-01-14".

# check data types
```{r}
glimpse(women)
```


Next Steps: Count years, and create a ggplot

# Transform date into years. See tutorial
```{r}

women <- women |> 
  mutate(year = year(published_date))

```


# Create a ggplot
```{r}
women |> 
  count(year) |> 
   ggplot(aes(x = year, y = n,
             fill = n)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
  scale_x_continuous(breaks=c(2015:2026)) + #fixes wonky x axis
  labs(title = "Years for Xinjiang Women Research",
       subtitle = "Limited NexisUni Run",
       caption = "Graphic by Rob Wells, 9-15-2026",
       y="Number of Calls",
       x="Year")



```



# fix date

```{r}
women <- women |> 
  mutate(date = as.Date(strptime(published_date, format = "%B %d, %Y %A")))
```