---
title: "Metadata"
author: "Wells"
date: "2026-09-15"
output: html_document
---

Class exercise to import your project metadata from the spreadsheet into R 

--Create a folder called "class_project" in your code folder 

--Save this file to the "class_project" folder 

--Move your Nexis spreadsheet to the folder 

--Create a project file on R Studio: File \| New Project \| \*\*\* select Existing folder \| "class_project" 

--After project.Rproj launches, then launch this file, metadata.Rmd, and complete the tutorial

**By doing this, you avoid the ring of hell where R can't find your data**

#load tidyverse, janitor, rio

```{r}

library(tidyverse)
library(janitor)
library(rio)
#install.packages("readxl")
library(readxl)

```

#Import your data

```{r}

data <- read_excel("NAME OF YOUR FILE.XLSX")
View(data)

```

# Clean names

```{r}
data <- data |> 
  clean_names()

```

# check data types
```{r}
glimpse(data)
```


Next Steps: Count years, and create a ggplot

# Transform date into years. See tutorial
```{r}

data <- data |> 
  mutate(year = year(published_date))

```


# Create a ggplot
```{r}
data |> 
  count(year) |> 
   ggplot(aes(x = year, y = n,
             fill = n)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
  scale_x_continuous(breaks=c(2015:2026)) + #fixes wonky x axis
  labs(title = "Years for Xinjiang Women Research",
       subtitle = "Limited NexisUni Run",
       caption = "Graphic by Rob Wells, 9-15-2026",
       y="Number of Calls",
       x="Year")



```

#Count publications

```{r}

```


#Visualize publications

```{r}

```




# NOTES ON FIXING COMPLEX Date COLUMN WITH HOURS AND MINUTES

```{r}
women <- women |> 
  mutate(date = as.Date(strptime(published_date, format = "%B %d, %Y %A")))
```