---
title: "Text Compiler-Vasectomy example"
author: "Rob Wells"
date: "2025-02-26-2025"
output: html_document
---

#Jour 689 \### This code, largely written by the famous **Sean Mussenden**, takes separate text files and compiles them into a single dataframe for analysis.

1)  Create a folder Week5_Text_Compiler
2)  Copy your spreadsheet index and the folder of text files into Week5_Text_Compiler
3)  Create an .Rproj file for Week5_Text_Compiler

```{r}
library(tidyverse)
library(janitor)

#install.packages("striprtf")
library(striprtf)

```
# Reformat .RTF files
```{r}
# Load required packages

# Set the paths for your folders
input_folder <- "./vasectomy_texts/"  # Replace with your input folder path
output_folder <- "./vasectomy_raw_text/" # Replace with your output folder path

# Create output folder if it doesn't exist
if (!dir.exists(output_folder)) {
  dir.create(output_folder)
}

# Get a list of all .rtf files in the input folder
rtf_files <- list.files(path = input_folder, pattern = "\\.RTF$", full.names = TRUE)

# Convert each .rtf file to .txt
for (file in rtf_files) {
  # Extract the file name without extension
  file_name <- tools::file_path_sans_ext(basename(file))
  
  # Read the RTF content
  rtf_content <- read_rtf(file)
  
  # Create output file path
  output_file <- file.path(output_folder, paste0(file_name, ".txt"))
  
  # Write the content to a .txt file
  writeLines(rtf_content, output_file)
  
  # Print progress
  cat("Converted:", file, "to", output_file, "\n")
}

cat("Conversion complete!\n")
```



# Raw text compiler

```{r include=FALSE}
#This creates an index with the file path to the stories. And then it compiles the stories into a dataframe
#####################
# Begin SM Code #####
#####################

###
# List out text files that match pattern .txt, create DF
###

#Adjust thisline for your file name
files <- list.files("./vasectomy_raw_text", pattern="*.txt") %>% 
  as.data.frame() |> 
  rename(filename = 1) |> 
  #create an matching file name
  mutate(index = str_replace_all(filename, ".txt", "")) %>%
  mutate(index = tolower(gsub("[[:punct:][:space:]|]", "", index))) |> 
  mutate(index = tolower(index)) |> 
  mutate(index = str_sub(index, 1, 25)) |> 
  distinct(index, .keep_all = TRUE)


#Join the file list to the index

final_data <- rio::import("catherine_Results list for_vasectomy, vasectomies.XLSX") |> 
  clean_names() |> 
   #create an matching file name
  mutate(index = tolower(gsub("[[:punct:][:space:]|]", "", title))) |> 
  mutate(index = tolower(index)) |> 
  mutate(index = str_sub(index, 1, 25)) |> 
  distinct(index, .keep_all = TRUE)
```

### Check for duplicate entries
```{r}

final_data |> 
  count(title) |> 
  arrange(desc(n))
```



```{r}

final_index <- final_data |> 
  inner_join(files, c("index")) |> 
#you need the actual hard-coded path on this line below to the text
  mutate(filepath = paste0("./vasectomy_raw_text/", filename))

head(final_index)
```
#Fact Check

```{r}

anti_final_index <- final_data |> 
  anti_join(files, c("index"))

```

#Checking for duplicates
```{r}
final_index |> 
  count(title) |> 
  arrange(desc(n))

```



#Text compiler

```{r}
###
# Define function to loop through each text file 
###

create_article_text <- function(row_value) {
  
  #row_value is the single argument that is passed to the function
  # Take each row of the dataframe
  temp <- final_index %>%
    slice(row_value)
  
  # Store the filename for  use in constructing articles dataframe
  temp_filename <- temp$filename
  
  # Create a dataframe by reading in lines of a given textfile
  # Add a filename column 
  articles_df_temp <- read_lines(temp$filepath) %>%
    as_tibble() %>%
    mutate(filename = temp_filename)
  
  # Bind results to master articles_df
  # <<- returns to global environment
  articles_df <<- articles_df %>%
    bind_rows(articles_df_temp)
}

###
# Create elements needed to run function
###

# Create empty tibble to store results
articles_df <- tibble()
#running once to test
#create_article_text(2) 
# Create an array of numbers to loop through, from 1 to the number of rows in our index dataframe 
row_values <- 1:nrow(final_index)

###
# Execute function using lapply
# This loops through each row of the dataframe and append results to master file
###

lapply(row_values, create_article_text)

###
# Clean up articles_df and join to index dataframe
###

articles_df <- articles_df %>%
  select(filename, sentence=value) %>%
  inner_join(final_index)


# write.csv(articles_df, "../exercises/assets/extracted_text/kemi_df2.csv")

```

# Notes Below

### Previous code to compile an index and add txt extensions

```{r}
#This tutorial shows how to copy files from one directory to another
# https://stackoverflow.com/questions/68995687/r-move-files-to-folder-based-on-list-or-column

inputdir  <- "/Users/robwells/Code/hcij_lynching_phase_two/narratives/data/articles_june_22_2024" 

targetdir <- "../exercises/assets/extracted_text" 
```

```{r}

#List out files : Sample code from previous use
# files <- list.files("~/Code/hcij_lynching_phase_two/narratives/code/articles_oct_19", pattern="*.txt") %>% 
#   as.data.frame() %>%
#   rename(filename = 1) 
#   filter(!str_detect(filename,"_bak"))
jackindex_oct19  <- jackindex_oct19  %>% 
   mutate(filename = paste(file_id2, ".txt", sep = ''))

#df <- extracted_articles_aug25$filename 

#creates new directory with 6448 articles: code/latest_articles/
filestocopy <- list.files(inputdir, full.names = TRUE) %>% 
  as_tibble() %>% 
  mutate(filename=(sub(".*/", "", value))) %>% 
  inner_join(jackindex_oct19, by=c("filename")) 
  #create a new column mutate = string-extract. keep everything after the last slash
  #inner_join to extracted articles aug 25
vector <- filestocopy$value
  #take that file path, select only filepath column and turn into a vector
  
filetocopy <- filestocopy[[1]]
#df_single <- df[[1]]
file.copy(from = filetocopy, to = targetdir)
map(vector, ~file.copy(from = inputdir, to=targetdir))

#map(filestocopy, ~file.copy(from = ., to=targetdir))


# Error: The size of the connection buffer (131072) was not large enough
# to fit a complete line:
# * Increase it by setting `Sys.setenv("VROOM_CONNECTION_SIZE")`
#https://github.com/tidyverse/vroom/issues/364
#Sys.setenv(VROOM_CONNECTION_SIZE = 10000000)
```
