---
title: "week3-KEY-homework-basic-r"
author: "Rob Wells"
date: "2024-09-06"
output: html_document
---

# Loading and Analyzing Data Exercise

This class exercise is a companion to the following chapter in Computational Text Analysis: [Ch. 11  Loading and Analyzing Data](https://wellsdata.github.io/CompText_Jour/docs/r-load-analyze-visualize.html)


#1: load tidyverse, rio and janitor libraries
```{r}
library(tidyverse)
library(janitor)
library(rio)

```


#2: Import Data:

https://github.com/wellsdata/CompText_Jour/raw/main/data/mob_analysis_data.csv
```{r}

mob <- read.csv("https://github.com/wellsdata/CompText_Jour/raw/main/data/mob_analysis_data.csv")

```

#3: Provide dimensions: Gives number rows, then columns

```{r}

dim(mob)
nrow(mob)
ncol(mob)
#RSW comment - Write that the Data has 3836 rows and 7 columns.
```

#4: Show names of your columns

```{r}
names(mob)
```

#5: Check data types

```{r}
glimpse(mob)

#rsw comment - correct or use str
```

#6: Display the first six rows

```{r}
head(mob)
```


#7: Create a table counting the number of entries by year and display the top 5 results

```{r}

mob %>% 
  group_by(Year) %>% 
  summarize(count = n()) %>%
  slice_max(order_by = count, n = 5)


```

#8: Create a table that counts the different entries in the "keyword" column

```{r}

mob %>% 
  group_by(keyword) %>% 
  summarize(count = n()) %>% 
  arrange(desc(count))


```


Create a simple column ggplot chart that shows the total entries by year

```{r}
mob %>% 
  group_by(Year) %>% 
  summarize(count = n()) %>% 
ggplot(aes(x = Year, y = count, fill = count)) +
  geom_col() +
  labs(title = "Count of Entries by Year", x = "Year", y = "Total Entries") +
  theme_minimal()

```

