---
title: "Bigrams Exercise"
author: "Rob Wells"
date: '2025-02-19'
output: html_document
---

# Jour 689 Spring 2025:

In this exercise, you will import the Nexis spreadsheet and create bigrams from the headlines

Setup. These instructions are important. Follow them carefully

1) Create a new folder called "bigrams_week_5" in your class folder
2) Copy this file to "bigrams_week_5"
3) Copy your spreadsheet to "bigrams_week_5"
4) Create an .Rproj file that points to "bigrams_week_5"

```{r}
#load tidyverse, tidytext, rio,, janitor libraries

library(tidyverse)
library(tidytext)
library(rio)
library(janitor)

```

```{r}
#Import dataframe 


bmore <- rio::import("Results list for__crime_ _Baltimore_.XLSX") |> 
  clean_names()

```
# Tokenize

```{r}

stories <- str_replace_all(bmore$hlead, "- ", "")
stories_df <- tibble(stories,)

# unnest includes lower, punct removal

stories_tokenized <- stories_df %>%
  unnest_tokens(word,stories)

stories_tokenized
```


#Remove stopwords


#Remove stopwords
```{r}

data(stop_words)

stories_tokenized <- stories_tokenized %>%
  anti_join(stop_words, by = c("word" = "word")) %>%
  filter(word != "temp_file") %>%
  filter(word != "stories_corpus") %>%
  filter(!grepl('[0-9]', word))

# Word Count

story_word_ct <- stories_tokenized %>%
  count(word, sort=TRUE)

```

# Bigrams

```{r}
bigrams <- stories_df %>%
  unnest_tokens(bigram, stories, token="ngrams", n=2)

bigrams

#Filter out stop words.


bigrams1 <- bigrams %>%
  separate(bigram, c("word1", "word2"), sep = " ")

bigrams2 <- bigrams1 %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)

bigram3 <- bigrams2 %>%
  count(word1, word2, sort = TRUE)


```
